Dependent Types and Martin-L\"of Type Theory: An Introduction

### 2016-08-05 paper.tex
more

\section{Exercises}
We have now seen and discussed all the types necessary to start proving more
complex proofs. As an exercise to the reader, we include a list of propositions
to prove. We invite the reader to use only the appropriate induction and
recursion principles in each proof. It helps to first attempt to prove these
propositions using simple pattern matching. As a hint, use a recursion principle
when the proof behaves like a simple function (i.e., defining {\tt `plus`}) and
induction principle when proving propositions {\tt C} for all terms of a given
type (i.e., like {\tt sym}).

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{code}
-- Identity Types
congruence : {A : Set} {x y : A} (P : A → Set) →
             (p : x ≡ y) → P x → P y
transitivity : {A : Set} {x y z : A} → (x ≡ y) →
               (y ≡ z) → (x ≡ z)

-- Natural numbers (requires the above)
_plus_ : ℕ → ℕ → ℕ
plus-left-unit : (i : ℕ) → 0 plus i ≡ i
plus-right-unit : (i : ℕ) → i plus 0 ≡ i
plus-successor : (i j : ℕ) → (suc i) plus j ≡ i plus (suc j)
plus-commutativity : (i j : ℕ) → i plus j ≡ j plus i
plus-associativity : (x y z : ℕ) →
                     x plus (y plus z) ≡ (x plus y) plus z

-- Lists
_++_ : {A : Set} → List A → List A → List A
reverse : {A : Set} → List A → List A
append-left-unit : {A : Set} (xs : List A) → Nil ++ xs ≡ xs
append-right-unit : {A : Set} (xs : List A) → xs ++ Nil ≡ xs
append-associativity : {A : Set} (xs ys zs : List A) →
                       xs ++ (ys ++ zs) ≡ (xs ++ ys) ++ zs
append-distributivity : {A : Set} (xs ys : List A) →
                        reverse (xs ++ ys) ≡ (reverse ys) ++ (reverse xs) 
double-reverse : {A : Set} (xs : List A) → reverse (reverse xs) ≡ xs

\end{code}
\end{minipage}
\end{center}

>>>>

\section{Exercises}
We have now seen and discussed all the types necessary to start proving more
complex proofs. As an exercise to the reader, we include a list of propositions
to prove. We invite the reader to use only the appropriate induction and
recursion principles in each proof. It helps to first attempt to prove these
propositions using simple pattern matching. As a hint, use a recursion principle
when the proof behaves like a simple function (i.e., defining {\tt `plus`}) and
induction principle when proving propositions {\tt C} for all terms of a given
type (i.e., like {\tt sym}). As an example, the first proof of each set of
proofs is provided:

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{code}
-- Identity Types
congruence : {A : Set} {x y : A} (P : A → Set) →
             (p : x ≡ y) → P x → P y
congruence P = 
  J (λ {x y} x=y → P x → P y)
    (λ x px → px)

transitivity : {A : Set} {x y z : A} → (x ≡ y) →
               (y ≡ z) → (x ≡ z)
\end{code}
\end{minipage}
\end{center}

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{code}
-- Natural numbers (requires the above)
_plus_ : ℕ → ℕ → ℕ
m plus n = f m n where
  f = recℕ (ℕ → ℕ) (λ n → n) (λ m r n → suc (r n))

plus-left-unit : (i : ℕ) → 0 plus i ≡ i
plus-left-unit = 
  indℕ (λ i → 0 plus i ≡ i)
        (refl zero)
        (λ x _ → refl (suc x))

plus-right-unit : (i : ℕ) → i plus 0 ≡ i
plus-successor : (i j : ℕ) → (suc i) plus j ≡ i plus (suc j)
plus-commutativity : (i j : ℕ) → i plus j ≡ j plus i
plus-associativity : (x y z : ℕ) →
                     x plus (y plus z) ≡ (x plus y) plus z
\end{code}
\end{minipage}
\end{center}

\begin{center}
\begin{minipage}{0.9\textwidth}
\begin{code}
-- Lists
_++_ : {A : Set} → List A → List A → List A
_++_ = append where
  append : {A : Set} → List A → List A → List A
  append {A} = recList (List A → List A)
                       (λ ys → ys)
                       (λ a xs f ys → a :: (f ys))

reverse : {A : Set} → List A → List A
append-left-unit : {A : Set} (xs : List A) → Nil ++ xs ≡ xs
append-right-unit : {A : Set} (xs : List A) → xs ++ Nil ≡ xs
append-associativity : {A : Set} (xs ys zs : List A) →
                       xs ++ (ys ++ zs) ≡ (xs ++ ys) ++ zs
append-distributivity : {A : Set} (xs ys : List A) →
                        reverse (xs ++ ys) ≡ (reverse ys) ++ (reverse xs) 
double-reverse : {A : Set} (xs : List A) → reverse (reverse xs) ≡ xs
\end{code}
\end{minipage}
\end{center}
